---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
```


```{r}
collision_data <- read.csv('Chicago_collision_data.csv')
```

```{r}
head(collision_data)
```
```{r}
collision_data_fullname <- collision_data %>% 
  unite(full_name, c('Genus', 'Species'), sep = ' ') 
head(collision_data_fullname)
```

```{r}
trophic_level <- read.csv('unique_name.csv')
trophic_level <- trophic_level %>% 
  select(-X, -X.1) %>% 
  rename(full_name = x)

head(trophic_level)
```


```{r}
# merge the collision dataframe with the trophic-level dataframe
collision_full <- merge(collision_data_fullname, trophic_level, by = 'full_name')

head(collision_full)
```

```{r}
# we only care about omnivore or non-omnivore, so merge carnivore and herbivore into non-omnivore
collision_full <- collision_full %>% 
  mutate(trophic_level = ifelse(trophic_level == 'Omnivore', 'Omnivore', 'Non-Omnivore'))

# wrangle the data using group_by and summarise to get the frequency of collisions per species
frequency_collision <- collision_full %>%
  group_by(full_name) %>% 
  summarise(frequency_of_collision = n(), trophic_level = first(trophic_level)) 

head(frequency_collision)  
```


```{r}
# Overview of the data
frequency_collision %>% 
  ggplot(aes(x = frequency_of_collision, fill = trophic_level)) + geom_histogram()
```
```{r}
# data distribution is right-skewed, so try to normalize it
ggplot(frequency_collision, aes(x = log(frequency_of_collision), 
                                fill = trophic_level)) + geom_histogram() + facet_wrap(~trophic_level)
```


```{r}
# check equal variances
frequency_collision %>% 
  group_by(trophic_level) %>% 
  summarize(v = var(frequency_of_collision))
```


```{r}
# log-transformation
frequency_collision_log_transformed <- frequency_collision %>% 
  mutate(l_t = log(frequency_of_collision))
head(frequency_collision_log_transformed)
# t-test
t.test(l_t~trophic_level, frequency_collision_log_transformed)
```

After log-transformation, still not normal. So instead using a nonparametric test: permutation test

```{r}
mean_table <- frequency_collision %>% 
  group_by(trophic_level) %>% 
  summarize(mean(frequency_of_collision))

# omnivore - non_omnivore
diff_means_obs <- 974.1111 - 679.4219	
```


Null hypothesis: No difference in the frequency of bird-building collisions in omnivore species and nonomnivore species.
mean_df = 0
Alternative hypothesis: Frequency is lower for omnivore species. 
mean_df < 0 (omnivore - non_omnivore)

```{r}
# permutation test
permute_and_calculate_mean_diff <- function(){
  reshuffled <- frequency_collision
  reshuffled$frequency_of_collision <- sample(reshuffled$frequency_of_collision, size = nrow(frequency_collision), replace = F)
  
  mean_omnivores_permuted <- mean(reshuffled %>% filter(trophic_level == "Omnivore") %>% pull(frequency_of_collision))
  mean_nonomnivores_permuted <- mean(reshuffled %>% filter(trophic_level == "Non-Omnivore") %>% pull(frequency_of_collision))
  
  mean_diff_permuted <- mean_omnivores_permuted - mean_nonomnivores_permuted
  # test statistic after permutation
  
  return(mean_diff_permuted)
}

permute_and_calculate_mean_diff()
```
```{r}
n_sims <- 1000 # number of times to permute data to generated null distribution
test_stats <- c()

for (i in 1:n_sims){
  test_stats[i] <- permute_and_calculate_mean_diff()
}

ggplot() + geom_histogram(aes(x = test_stats), fill = "gray") +
  geom_vline(xintercept = diff_means_obs, color = "red")
```
```{r}
length(abs(test_stats)[(diff_means_obs > test_stats)])/length(test_stats)
```

Fail to reject the null hypothesis. 


```{r}
# sample size
frequency_collision %>% group_by(trophic_level) %>% summarize(count = n())
```









