---
title: "Bird-Building Collision Rate"
author: "Yunhua"
date: "2023-12-01"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data
```{r message=TRUE, warning=FALSE}
library(tidyverse)
library(stringr) 
bird<-read.csv(file = "merged_data.csv")
head(bird) #Show the first six lines of observations.
```

## 1.	Statistical Analysis

### Descriptive Statistics
```{r message=FALSE, warning=FALSE}
library(pastecs)
round(stat.desc(bird[-1]),4)##descriptive statistic
ggplot(bird,aes(x=Wind_Load))+geom_histogram()
ggplot(bird,aes(x=collision_freq))+geom_histogram()
```

Data distribution of the variable collision_freq showed a serious right bias distribution, indicating that there was a maximum value in the data, and logarithmic transformation should be used.

```{r}
bird<-bird %>% 
  mutate(lfreq = log(collision_freq)) ##Generate a new variable to logarithmize collision_freq.
ggplot(bird,aes(x=lfreq))+geom_histogram()
```

### Check normality
```{r}
transformed_data <- bird$lfreq
shapiro_test_result <- shapiro.test(transformed_data)
print(shapiro_test_result)
```

###	Correlation Analysis

```{r}
cor.test(bird$Wind_Load,bird$lfreq,method = "pearson") #correlation coefficient
pairs(bird[,2:6])
```
The relationship between variables is shown in the matrix scatter plot. `wind load` and `lfreq` have a weak positive correlation.

## Linear Mixed Models

Because the multiple linear regression model has some limitations, this study adds the date variable to create a linear mixed model, explores the impact of wind load on the collision rate on different dates, and compares the explanatory power and goodness of fit between the linear regression model and the linear mixed model.

#### Merge data
```{r}
data1<-read.csv(file = "Chicago_collision_data.csv")
data1$Species_Name<-str_c(data1$Genus,data1$Species, sep = " ")
bird1<-merge(data1,bird,by="Species_Name")
```

#### Building model 
```{r message=FALSE, warning=FALSE}
library(lme4)
library(lmerTest)
fit2<-lmer(lfreq~Wind_Load+(1|Date),data = bird1)
summary(fit2)
```

```{r}
# Create a new dataframe with the fitted values
bird1_fitted <- data.frame(Wind_Load = bird1$Wind_Load, Fitted_Values = fitted(fit2))

# Plotting
ggplot(bird1_fitted, aes(x = Wind_Load, y = Fitted_Values)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Wind Load", y = "Predicted Frequency", title = "LMM: Wind Load vs Predicted Frequency")
```

