---
title: "Prediction 1"
---

Reading in and examining the data.  
```{r}
library(tidyverse)
library(dplyr)
bird_data <- read.csv("Chicago_collision_data.csv")
light_data <- read.csv("Light_levels_dryad.csv")

head(bird_data)
tail(bird_data)
str(bird_data)
glimpse(bird_data)

bird_data %>% 
  count(Species)

# There are different numbers of observations on each date. 
bird_data %>% 
  group_by(Date) %>% 
  tally()

# There are 36,315 observations at McCormick place (MP) and 33,469 observations 
# at other locations in Chicago (CHI). Observations at the CHI locations were
# collected by a volunteer network. Exclude? Note on page 3 of primary article, 
# authors state in section 1c) that exploratory analysis yielded similar results
# when analysing MP separately from rest of Chicago (CHI). Use this to justify
# excluding CHI observations if needed? Third justification is that light_data
# is only at McCormick place, so it doesn't make sense to extrapolate this to
# other locations across the city that could be a lot darker or brighter. 

bird_data %>% 
  count(Locality)

# The bird collision data goes from October 1978 to November 2016. 
# Primary article says that bird building collisions were observed daily during
# spring (March to May) and autumn (August to November). 
bird_data %>% 
  reframe(range(Date))

head(light_data)
tail(light_data)
str(light_data)
glimpse(light_data)

# The light intensity data ranges from a value of 3 to 17. 
# From primary article, "Light levels are the sum of the number of lighted
# windows during pre-dawn surveys from 2000 to 2018, ranging as integers from
# 1 to 17. 
light_data %>% 
  reframe(range(Light_Score))

light_data %>% 
  reframe(range(Date))

# Note that light_data is nighttime light from McCormick place, and bird_data is
# observations of fatal bird collisions in the *mornings*. 

```

Modifying `bird_data` to count data to get total number of collisions per night.
Will also need to filter dates from 2000 to 2018 because this is the range
covered by `light_data`. But `bird_data` only goes to 2016, so filter from 2000
to 2016. 

```{r}
# Filter out MP only based on justifications above. 
bird_data_MP <- bird_data %>% 
  filter(Locality == "MP")

# In order to filter for observations between 2000 and 2018, filter by row from 
# where date becomes 2000 to where date is 2018. 2000 starts at row 36 (date
# is 2000-05-14) and 2018 ends at...

bird_data_MP

# But the collisions are presented in chunks by species! Ammodramus nelsoni changes
# to Ammodramus henslowii at row 52! 
# So the data is not ordered by date... And the dates are different for different
# species (i.e. the same set of dates for Ammodramus nelsoni does not repeat for
# Ammodramus henslowii). 

bird_data_MP %>% 
  count(Species)
# There are 83 species... 

# Side note - there are a lot more observations for some species than others.
# So total collisions by date could be affected by greater local abundance
# of some species (not just light intensity)... date as random
# effect?

dates <- bird_data_MP %>% 
  group_by(Date) %>% 
  tally()

# A data frame with Dates in one column and number of collisions n in another 
# column. 4437 rows, 2 columns. 

dates %>% 
  reframe(range(Date))

# Confirming that this df currently goes from 1978 to 2016. 

# Year 2000 starts at row 2734 (found this by subsetting and checking different 
# rows of dates). 

dates[2734,]

dates_after_2000 <- dates[2734:nrow(dates),]

tail(dates_after_2000)

dates_collisions <- dates_after_2000 %>% 
  rename("Total_collisions" = "n")

```


Modifying `light_data` to include a column with total number of collisions. This
will be the data frame that is used for analysis. 

```{r}
# Filter to 2000-2016 range. light_data is 3067 rows. 2016 ends at row 2829. 

light_data[2829,]

tail(light_data)

light_range <- light_data[1:2829, ]

tail(light_range)

light_range <- light_range %>% 
  rename("Light_intensity" = "Light_Score")


# So now how to add a column to dates_collisions? dates_collisions is 1704 rows,
# light_range is 2829 rows. 

# merge has an argument all.x whose default is FALSE, so rows which do not have 
# a match are automatically excluded. 

data <- merge(dates_collisions, light_range)

head(data)
tail(data)

library(lubridate)
data2 <- data
data2$Date <- as_date(data2$Date)

head(data2)

```


Checking assumptions of a linear model. Note that authors used a GLM to 
investigate effect of flight calls on collision (collision index). 
How is light distributed, and how is collision frequency distributed? 
What kind of data is it? 

```{r}


library(ggplot2)

# 1) Are data normally distributed? 
# Bird-building collisions are not normally distributed. 
data %>% 
  ggplot(aes(x = Total_collisions)) +
  geom_histogram(binwidth = 2.5, fill = "blue", colour = "black") + 
  labs(x = "Total bird-building collisions per day",
       y = "Count")

# Log or sqrt transforming doesn't work. 

data %>% 
  ggplot(aes(x = log(Total_collisions))) +
  geom_histogram()

data %>% 
  ggplot(aes(x = sqrt(Total_collisions))) +
  geom_histogram()

# Using Poisson distribution. Count data, and in theory independent (bird 
# colliding with a building does not affect probability of another bird colliding 
# with building).  

# 2) Homogeneity of error variances
lm_test <- lm(Total_collisions~Light_intensity, data = data)
summary(lm_test)

par(mfrow = c(2,2))
plot(lm_test)

# This assumption is clearly violated. 

lm_test_log <- lm(log(Total_collisions)~Light_intensity, data = data)
summary(lm_test)

par(mfrow = c(2,2))
plot(lm_test_log)

# But actually slightly better with log transform. GLM log link?  


```

So can't use a simple linear model. Going to try 1) GLM and 2) LMM. 
Predicting that LMM is better to use because suspect that data is not independent (factors on 
a particular date that might affect total collisions).

```{r}
model_glm <- glm(Total_collisions~Light_intensity,
                 family = poisson,
                 data = data2)

summary(model_glm)


data %>% 
  ggplot(aes(x = Light_intensity,
             y = Total_collisions)) +
  geom_point(position = "jitter") +
  geom_smooth(method = "glm",
              method.args = list(family = "poisson")) +
  labs(x = "Light intensity",
       y = "Total bird-building collisions per day")

# Trying a LM with date. 
# This takes a while, because it includes a term for each date (we have 1641 rows 
# i.e. dates when collisions are observed, in the data frame!). 

model_lm <- lm(Total_collisions~Light_intensity + Date,
                data = data)

summary(model_lm)

# But can't use this one, because it's a linear model (and the collision data
# violates the assumptions of LMs). 

# Going to try a LMM random intercept model with date as a random effect.(2nd model) 
library(lme4)
library(lmerTest)

model_lmm <- lmer(Total_collisions~Light_intensity + (1|Date),
                  data = data2,
                  REML = FALSE)

summary(model_lmm)

data_with_lmm <- data2 %>% 
  mutate(fit_lmm = predict(model_lmm))

data_with_lmm %>% 
  ggplot(aes(x = Light_intensity,
             y = Total_collisions,
             colour = Date)) + # could colour by date, but this would be very difficult to visualise!
  geom_point(position = "jitter", alpha = 0.8) +
  geom_abline(aes(intercept = 2.2230, slope = 0.3097), colour = "blue", linewidth = 1) +
  labs(x = "Light intensity",
       y = "Total bird-building collisions per day")


AIC(model_glm, model_lmm, k = 2)

```

Indeed, the LMM is better. 

So... (reporting and interpreting results)

```{r}
summary(model_lmm)
```

The model finds that light intensity has a significant effect on the number of 
bird-building collisions (p-value 1.01e-11). Increasing light intensity by 1 unit
increases the number of bird-building collisions by 0.3097. 

Date accounts for 79.815 / (79.815 + 3.172) * 100 = 96.17 ~ 96% of the residual variance (the variance leftover after accounting for light intensity as a fixed effect). 






